---
title: "Coursework_1"
output:
  html_document:
    df_print: paged
---

```{r}
# loading all the packages I need for the questions
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(readxl)
library(RColorBrewer)
library(scales)
```

# Question 1

## a
```{r}
read_lines(here("data_cw", "data_raw","nasa_global_temperature.txt"),n_max = 10) # having a look at the first 10 lines to see what I need to do
nasa_temp <- read_table(here("data_cw", "data_raw","nasa_global_temperature.txt"),skip = 3) %>% # skipping the unnecessary parts 
  remove_missing() %>% # removing anything that is unnecessary 
  rename(date = 'Year', temp = No_Smoothing) %>% #renaming variables 
  mutate(date = parse_date(date,format = "%Y")) %>% #changing the type of variable to date type
  select(date, temp) #only leaving the variables I want 
  
nasa_temp # table 

  nasa_temp_plot <- ggplot(nasa_temp, mapping = aes(x = date, y = temp)) + # creating the plot and what to put on the axes
    geom_point() + # using a scatterplot graph
    ggtitle("Arctic temperatures", "Temperatures of the arctic from 1880 to 2020") + # placing a suitable title 
    ylab("Temp (Degrees Celsius)") + # labelling the axes
    xlab("Year")

  nasa_temp_plot # visualisation 
```

## b

```{r}
read_lines(here("data_cw", "data_raw","nasa_arctic_sea_ice.csv"),n_max = 10) # having a look at the first 10 lines to see what I need to do
nasa_ice <- read.csv2(here("data_cw", "data_raw","nasa_arctic_sea_ice.csv")) %>% # read csv2 because it uses a different delim 
  rename(date = 'year', ice = extent) %>% #renaming variables 
  mutate(date = as.character(date)) %>%  #changing the column types have to turn to character to parse as date 
  mutate(date = parse_date(date, format = "%Y")) %>%
  select(date, ice) # selecting the variables I want

nasa_ice

nasa_ice_plot <- ggplot(nasa_ice, mapping = aes(x = date, y = ice)) + # creating the plot and what to put on the axes
  geom_smooth() + # I thought this type of graph would be best 
  ggtitle(" Arctic ice levels", "Levels of ice in the arctic from 1979 to 2020") + # adding labels to my visualisation
  ylab("sea ice (per million sq km)") +
  xlab("Year")



nasa_ice_plot
  
```
## c 

```{r}
read_lines(here("data_cw", "data_raw","nasa_sea_level.csv"), n_max = 50) # having a look at the first 50 lines to see what I need to do 
nasa_sea <- read_log(here("data_cw", "data_raw","nasa_sea_level.csv"),skip = 48) %>% # the best function to read the file
  remove_missing() %>% # removing anything unnecessary
  select(X3, X12) %>% # selecting the columns I want 
  rename(date = X3, sea = X12) %>% # giving the columns the appropriate names 
  mutate(date = date_decimal(date)) %>% # changing the column type 
  mutate(date = as_date(date))

nasa_sea

nasa_sea_plot <- ggplot(nasa_sea, mapping = aes(x = date, y = sea)) + #creating the plot and what to put on the axes
  geom_line() +
  ggtitle("Sea levels", "Change in sea levels over time") +
  xlab("Year") +
  ylab("Sea levels (mm)")
  
nasa_sea_plot
  


```

## d

```{r}
read_lines(here("data_cw", "data_raw","nasa_carbon_dioxide.txt"), n_max = 60) # looking at the lines to see what to do 

nasa_co2 <- read_table(here("data_cw", "data_raw","nasa_carbon_dioxide.txt"), skip = 52) %>%
  select("#", date, alized) %>% # selecting the variables I want 
  unite(date, "#", date, sep = "") %>% # uniting two columns together to make one column which is the date 
  rename("co2" = "alized") %>% # renaming variables
  mutate(date = parse_date(date, format = "%Y%m")) # turning column into a date type

nasa_co2
         
nasa_co2_plot<- ggplot(nasa_co2, mapping =  aes(x = date, y = co2)) + # creating a plot with labels 
  geom_col() +
  coord_cartesian(ylim = c(300,450)) + # there were two data points that were extreme outliers so they were removed from the plot 
  ggtitle("Co2 levels over time", ("Average global co2 level from 1958 to 2020")) +
  xlab("") +
  ylab("Co2 level (ppm)")

nasa_co2_plot


```

## e

```{r}

nasa1 <- full_join( nasa_co2, nasa_temp, by = "date") %>% # joined two tibbles together 
  arrange(date)

nasa2 <- full_join(nasa_ice, nasa_sea, by = "date") %>% # joined two tibbles together 
  arrange(date)
  
nasa <- left_join(nasa1, nasa2) %>% #joined all the tibbles together 
  arrange(date)

nasa

```

## f

```{r}
nasa_plot<- nasa %>% 
  select(date, temp, co2) %>%
  filter(date >= as.Date("1960-01-01")) #this is to get the years past 1960 and between 2020

temp_co2 <- ggplot(nasa_plot, mapping = aes(x = temp, y = co2, colour = date)) +
  geom_point(position = "jitter") + # this is to prevent overplotting
  ggtitle("Temperature and Co2 levels from 1960") + # adding titles and labels 
  ylim(300, 400) + # to remove the outliers from the graph
  xlab("Temp (Degrees celsius)") +
  ylab("Co2 levels") +
  theme(panel.grid.major.x = element_line(color = "grey90"), # this is to make the graph better visually and to see the colour scheme better by removinh the backround etc
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(colour = "grey90"), 
    panel.grid.minor.y = element_blank(),
    panel.background = element_blank(),
    axis.line.x = element_line("black"),
    axis.line.y = element_line("black")
  )

temp_co2 + 
  labs(colour = "Year") # labeling the legend 
```



# Question 2

## a

```{r}
read_lines(here("data_cw", "data_raw","luthi_carbon_dioxide.txt"), n_max = 784) # reading lines to know what I am doing 

historic_co2 <- read_table(here("data_cw", "data_raw","luthi_carbon_dioxide.txt"), skip = 773) %>% # turning into tibble
  rename(yrbp = "Age(yrBP)", co2 = "CO2(ppmv)") # renaming variables
  
  
```

## b

```{r}

historic_co2 <- historic_co2 %>%
  mutate(yrbp = yrbp + 13) %>% # thirteen years from 2008 makes 2021 which is the present
  select(yrbp, co2) # selecting the variables I want 
 
nasa_co2_1 <- nasa_co2 %>%
   mutate(yrbp = as.duration(((today()) - (date)))/dyears(1)) %>% # to get the years before present 
   mutate(yrbp = round(yrbp)) %>% # rounding the values to create whole numbers
   select(yrbp, co2) # selecting the variables I want

 nasa_co2_1
 historic_co2

combined_co2 <- full_join(nasa_co2_1, historic_co2, by = c("yrbp", "co2"))  # this is to create the combined_co2 tibble needed for the graph

combined_co2

``` 

## c

```{r}

combined_co2_plot <- ggplot(combined_co2, mapping = aes(x = yrbp, y= co2)) + #creating a plot with labels
  geom_line() + # using the same graph as the figure
  ylim(180, 400) + # setting the limits to recreate as accurately as possible
  xlab("Years before present") + # creating the same labels
  ylab("Carbon dioxide [ppm]") + 
   theme(panel.grid.major.x = element_blank(), # creating a theme to recreate the graph
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(), 
    panel.grid.minor.y = element_blank(),
    panel.background = element_blank(),
    axis.line.x = element_line("black"),
    axis.line.y = element_line("black")
  )

combined_co2_plot <- combined_co2_plot + # changing the font sizes to recreate the graph
  theme(
  axis.title.x = element_text(size = 18),
  axis.text.x = element_text(size = 12),
  axis.title.y = element_text(size = 18),
  axis.text.y = element_text(size = 12)
)

(combined_co2_plot <- combined_co2_plot + scale_x_reverse(labels = comma)) # add a comma as a grouping mark




```
## c continued

```{r}
combined_co2_plot <-combined_co2_plot + # this is to add the annotation however I could not quite find the colour 
annotate("text", x = 250000, y = 360, label = "2021 average:
417 ppm", colour = "red2")

combined_co2_plot <- combined_co2_plot + # adding the arrow
  annotate("curve", x = 150000, xend = 10000 , y = 350, yend = 400, colour = "red2", size = 1, arrow=arrow())

```



# Question 3

## a

```{r}

sea_ice_1 <- read_xlsx(here("data_cw", "data_raw","nsidc_sea_ice_daily_extent.xlsx")) %>% # reading data into tibble 
  rename("month" = "...1", "day" = "...2") %>% # renaming variables into the appropriate names
  fill(month) %>% # to ensure that months are all present and to prevent NAs
  mutate(day = as.integer(day), month = match(month, month.name)) %>% # turning year month and day into integers 
  select(-c("...47",contains("1981-2010"))) # selecting everything except unnecessary columns 
 
sea_ice_1_dates <- sea_ice_1 %>% # making dates 
  select(-c(month, day))

sea_ice_1_m_d <- sea_ice_1 %>% # making a tibble with the month and days as columns 
  select(month, day)

sea_ice_1_dates <- sea_ice_1_dates %>% # this is to get the dates 
  pivot_longer(everything(), names_to = "year")


sea_ice_1 <-  cbind(sea_ice_1_dates,sea_ice_1_m_d) # binding the column together 

sea_ice_1 <- sea_ice_1 %>%
  rename("extent" = "value") %>% 
  select(extent,everything()) %>% 
  mutate(day = as.integer(day), year = as.integer(year)) %>%
  group_by(month) 

sea_ice_1
```

## b

```{r}
b_avg <- sea_ice_1 %>% # creating a baseline
  filter(year==1979) %>%
  arrange(month,day) %>%
  unique() %>%
  group_by(month) %>%
  summarise("baseline"= mean(extent, na.rm = TRUE)) 
 
b_avg
   
m_avg <- sea_ice_1 %>% # creating monthly averages
  group_by(year, month) %>% 
  summarise("monthly_averages"= mean(extent, na.rm = TRUE)) 

m_avg

sea_ice_2 <- left_join(b_avg, m_avg) %>% # dividing baseline and monthly averages to create proportion_baseline_extent
  mutate("proportion_baseline_extent" = monthly_averages/baseline) %>%
  select(year, month, proportion_baseline_extent)
 
sea_ice_2%>%
  group_by(month)
 
```

## c

```{r}

sea_ice_2_plot <- sea_ice_2 %>% # creating the plot 
  filter(year > 1978 & year < 2021) %>% # filtering the appropriate years
  group_by(month) %>%
  ggplot(sea_ice_2, mapping = aes(y =month.name[month], x = year)) + # turning the month into month names 
  geom_raster(aes(fill = proportion_baseline_extent)) + 
  scale_fill_distiller(palette = "RdPu") # getting the colour 
  
sea_ice_2_plot <- sea_ice_2_plot + # using theme to recreate graph as accurately as I can 
  theme(panel.grid.major.x = element_blank(), 
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(), 
    panel.grid.minor.y = element_blank(),
    panel.background = element_blank(),
    axis.line.x = element_line("black"),
    axis.line.y = element_line("black")
  )

sea_ice_2_plot <- sea_ice_2_plot + # adding labels 
  ggtitle("Sea ice (northern hemisphere)") +
  xlab("Year") +
  ylab("Month") +
  labs(fill = "Proportion of 1979 extent")

sea_ice_2_plot

  
```


# question 4

## a
```{r}
read_lines(here("data_cw", "data_raw","stop_and_search.csv"), n_max = 10) #reading the first few lines 

stop_search_1 <- read_csv(here("data_cw", "data_raw","stop_and_search.csv")) %>% #reading into tibble 
  rename("stops" = "Number of stop and searches",  "population" = "Population by ethnicity", "rate" = "Rate per 1,000 population by ethnicity", "year" = "Time") %>%  # giving variables appropriate names 
  clean_names() %>% # tidying names 
  select(year, ethnicity, legislation_type, geography, stops, population, rate) %>% # selecting the columns I want to keep
  filter(ethnicity == "White" | ethnicity == "Black" | ethnicity == "All" | ethnicity == "Other" | ethnicity == "All") %>% # filtering ethnicities I want 
  mutate("population" = as.numeric(population, na.rm = TRUE), "rate" = as.numeric(rate, na.rm = TRUE) , "stops" = as.numeric(stops, na.rm = TRUE)) %>% # changing col types 
  mutate(year = year(as.Date(sub('[0-9]{2}-', '', year), format = '%Y'))) # turning year into date type parsing did not work so I decided to split the years up instead 



stop_search_1


```

## b
```{r}
white_rate <- stop_search_1 %>%
  filter(ethnicity ==  "White") %>% # filter ethnicity to white only
  remove_missing() %>% # remove missing parts
  select(rate) %>%
  rename("white_rate" = rate )# creating a white rate 

stop_search_rate <- stop_search_1 %>% 
  select(rate) %>%
  remove_missing()

 # white_rate/stop_search_rate




```

## c
1.**_What year had the highest number of stop and searches and what are the possible reasons behind this ?_** This is operationalised since we are looking at the highest number of searches furthermore looking we can use this to further look for potential reasons why.

2.**_Is there a correlation between the population and the amount of stop and searches?_** This is operationalised since we will be looking at the correlation between two variables.

3.**_Do the rate of stops by ethnicity differ from 2014 onwards?_** This will look at which ethnicity is being searched at a different rate in the most recent years this dataset has available. 

## d 

### Question 1

```{r}
yearly_stops <- stop_search_1 %>% #selecting and creating a plot to help me with question 1
  select(year, stops) %>%
  arrange(desc(stops)) %>%
  group_by(year) %>%
  summarise("total_stops" = sum(year)/100) %>%
  mutate(year= as.character(year)) %>%
  mutate(year = parse_date(year, format = "%Y")) 

yearly_stops


yearly_stops_plot <- ggplot(yearly_stops, mapping = aes(x = year, y = total_stops)) +
  geom_col() +
  scale_y_continuous(labels = comma) + 
  xlab("Year") +
  ylab("Stops (Per 1000 stops)") +
  ggtitle("Total amount of stops per year")

yearly_stops_plot
```
The graph shows that 2019 has had the most stop and searches performed which answers the initial question. However looking closer into the graph it raises new questions, the stops seemed to have increased by over 100,000 in 2019 compared to the other years leading to the question of why there has been such a steep increase in searches. 

### Question 2
```{r}
population_stop <- stop_search_1 %>% 
  select(population, stops) %>%
  remove_missing() %>%
  group_by(population) %>%
  summarise(total_stops = sum(stops)) %>%
  mutate(percentage_of_pop = (total_stops/population)*100) %>% # used to find the percentage
  mutate(population_per_mil = population/1000000) # population per mil to help with scaling the graph

population_stop

ggplot(population_stop, mapping = aes(y = total_stops , x = percentage_of_pop, colour = population_per_mil)) +
  geom_point(position = "jitter") + #this is to help with overplotting 
  scale_y_continuous(labels = comma) +
  ggtitle("Proportion of stops", " the percentage of stops for every population") + #adding labels 
  xlab("Percentage of population stopped") +
  ylab("Total stops")
 
```
This graph helps answer the question to a certain degree where the higher the population, the less total stops there were. And the higher percentage of stops were due to a lower population therefore less stops were made in total but there were a higher proportion of stops made. This raises new questions and theories such as the idea that the police can search smaller populations as it is less difficult, this graph however only has meaningful extremes, to support the answers of question 2 we need to zoom in on the cluster (see below)

```{r}
ggplot(population_stop, mapping = aes(y = total_stops , x = population_per_mil , colour = percentage_of_pop )) + # adding limits to sort of zoom in the graph
  geom_point(position = "jitter") +
  scale_y_continuous(labels = comma) +
  ggtitle("Proportion of stops", " the percentage of stops for every population") +
  xlab("Population per mil") +
  ylab("Total stops") +
  ylim(0,30000)+
  xlim(0,10)
```

When zooming in to the cluster, it shows that there is a lower percentage of the population getting stopped and the majority of the points are of a population of less than 1 million, showing that there is indeed a positive correlation between population and the number of stop and searches, but when looking closer there is a higher percentage of people getting searched when the population is lower. 

### question 3

```{r}
stop_search_1_2015_onwards <- stop_search_1 %>%
  filter(year>=2014) %>%
  select(ethnicity, geography, rate, stops, year) %>%
  group_by(geography) 
stop_search_1_2015_onwards 

ggplot(stop_search_1_2015_onwards, mapping = aes(x = ethnicity, y = rate)) +
  geom_col() +
  facet_wrap(~year) # faceting by years 



```

This shows that the black ethnicity have had the higher rate of stop searches compared to those of "white" and "other" ethnicity, this may suggest that black people are disproportionately stopped.
