---
title: "Practical_7"
output: html_notebook
---

7.0 Exercise
Start a new R notebook named practical_07.Rmd and save it to the rmd folder of your RStudio project practicals. Make sure that all the R packages you need are loaded at the beginning. All the following exercises should be attempted within R code chunks of this notebook.

```{r}
library(tidyverse)
library(ggrepel)
library(janitor)
library(lubridate)
library(gapminder)
```

# 7.1 Exercise
Consider again the gapminder data discussed in the lecture notes (recall that this data set is made available through library("gapminder")).
```{r}
gapminder
```


a. Create the following figures.

i. a histogram of life expectancies in different countries for the year 2007;

```{r}
gapminder_2007_plot <- gapminder %>%
  filter(year == 2007) %>%
  group_by(country) %>%
  ggplot(mapping = aes(x = lifeExp))

gapminder_2007_plot + geom_histogram()

gapminder_2007_plot
```


a boxplot of life expectancies in different countries for the year 2007;

```{r}
gapminder_2007_plot + geom_boxplot()
# I think I need the other grouping function
```

iii. a line graph showing how life expectancy has evolved in the United Kingdom;

```{r}
gapminder %>%
  filter(country == "United Kingdom") %>%
  ggplot() +
  geom_line(mapping = aes(y = lifeExp, x = year))
```


iv. a scatterplot of GDP per capita vs life expectancy.

```{r}
gapminder %>%
  ggplot() +
  geom_point(mapping = aes(x = lifeExp, y = gdpPercap))
```


b. Recreate the following figure which shows the GDP per capita for countries in Asia in 2007.

[image](~/Downloads/image_ggplot_2.png)


```{r}
gapminder_asia_plot <- gapminder %>%
  filter(continent == "Asia" & year == "2007") %>%
  ggplot() +
  geom_col(mapping = aes(x = gdpPercap, y = country))
```

c. Unless your data have a natural order (as, e.g., in time-series data), it is often sensible to order your data. With this in mind, improve the following figure. Hint: you may find the base R function reorder() useful which takes two arguments and which thus allows you to reorder the entries in x according to the elements in y by typing reorder(x, y).

```{r}
gapminder_asia_plot+
  geom_col(mapping = aes(y = reorder(country, gdpPercap), x = gdpPercap, fill = lifeExp))

```

d. Visualise, using a single line, how the global average life expectancy has evolved between 1952 and 2007. Make sure to take the different population sizes in different countries into account. You may find the function weighted.mean() helpful.

```{r}
global_life_exp <- gapminder %>%
  filter(between(year,1952, 2007)) %>%
  group_by(year) %>%
  summarise(weighted_avg_life_exp = weighted.mean(lifeExp, w = pop))
# assuming w stands for weighting 

(global_life_exp_plot <- global_life_exp %>%
  ggplot() +
  geom_line(mapping = aes(x = year, y = weighted_avg_life_exp)))
```


e. Visualise how the life expectancy in your own country has evolved relative to the weighted average of other, similar countries, e.g. its neighbours (with weighting by population size).

```{r}
global_life_exp <- gapminder %>%
  filter(between(year,1952, 2007)) %>%
  group_by(year) %>%
  summarise(weighted_avg_life_exp = weighted.mean(lifeExp))

ghana_life_exp <- gapminder %>%
  filter(between(year,1952, 2007),
         country == "Ghana"
  ) %>%
  group_by(year) %>%
  summarise(ghana_avg = weighted.mean(lifeExp))

relative_gh <- left_join(global_life_exp, ghana_life_exp, by = "year")

relative_gh %>%
  ggplot() +
  geom_line(mapping = aes(x = year, y = weighted_avg_life_exp), colour = "green") +
  geom_line(mapping = aes(x = year, y = ghana_avg))
```

f. Visualise the how the life expectancy has evolved between 1952 and 2007 for all countries in the data set without differentiating them visually (i.e. draw one black line for each country).

```{r}
 # have to use the group thing 

global_life_exp_2 <- gapminder %>%
  filter(between(year,1952, 2007)) %>%
  group_by(year, country) %>%
  summarise(weighted_avg_life_exp = weighted.mean(lifeExp))

global_life_exp_2

global_life_exp_plot <- global_life_exp_2 %>%
  ggplot() +
  geom_line(mapping = aes(x = year, y = weighted_avg_life_exp, group = country))

global_life_exp_plot

# damn 

```


g. Create the same graph as before but now ensure that the lines have different colours depending on the continent to which a country belongs. Also plot the same figure but with different line types instead of colours for each continent.

```{r}
# different colours

global_life_exp_3 <- gapminder %>%
  filter(between(year,1952, 2007)) %>%
  group_by(year, country, continent) %>%
  summarise(weighted_avg_life_exp = weighted.mean(lifeExp))

global_life_exp_3

global_life_exp_plot <- global_life_exp_3 %>%
  ggplot() +
  geom_line(mapping = aes(x = year, y = weighted_avg_life_exp, colour = continent, group = country))

global_life_exp_plot

# different linetype 

global_life_exp_plot_2 <- global_life_exp_3 %>%
  ggplot() +
  geom_line(mapping = aes(x = year, y = weighted_avg_life_exp, linetype = continent, group = country))

global_life_exp_plot_2

```


h. Visualise how the continent-wise average life expectancy has evolved between 1952 and 2007 via a suitable graph.

```{r}
gapminder_continent <- gapminder %>%
  filter(between(year, 1952, 2007)) %>%
  group_by(year, continent) %>%
  select(continent, year, lifeExp) %>%
  mutate(avg_life_exp = mean(lifeExp))

gapminder_continent

gapminder_continent %>%
  ggplot() +
  geom_line(mapping = aes(x = year, y = avg_life_exp, group = continent, colour = continent))
```

i. Create a scatterplot of the base-10 logarithm of GDP per capita vs life expectancy in 2007 such that the colour of the points differs with continent and the size of the points gives information about the population size.

```{r}
gapminder_2007 <- gapminder %>%
  filter(year == 2007)

gapminder_2007 %>%
  ggplot() +
  geom_point(mapping = aes(x = log10(gdpPercap), y = lifeExp, colour = continent, size = pop))
```


j. Plot a scatterplot of the base-10 logarithm of GDP per capita vs life expectancy such that the colour of the points is differs with the year, i.e. colour = year. What happens if you map colour = as.factor(year) or colour = as.ordered(year)instead. Which do you find more sensible?

```{r}
   p <- ggplot(
           data    = gapminder, 
           mapping = aes(x = log10(gdpPercap), y = lifeExp)
         )
    
    # continuous colour scale doesn't seem appropriate because 
    # we only have round years in our data set.
    p + geom_point(mapping = aes(colour = year)) 
    
    # qualitative colour scheme doesn't seem appropriate either because we lose 
    # all the information about the ordering (in time).
    p + geom_point(mapping = aes(colour = as.factor(year)))
    
    # sequential but discrete colour scheme seems more appropriate.
    p + geom_point(mapping = aes(colour = as.ordered(year)))
```


k. Explain what is wrong with the following (without running the code).
```{r}
ggplot(data = gapminder) +
  geom_point(mapping = aes(x = log10(gdpPercap), y = lifeExp, shape = pop))
```

We are trying to map a continuous variable to a shape which is not possible as shapes only work with discrete variables.

l. ggplot(data = gapminder) +
  geom_point(mapping = aes(x = log10(gdpPercap), y = lifeExp, shape = pop))
Reproduce the following figures in which pop_in_bn is the total global population [in billion persons] and gdp_in_tn is the total global GDP [in trillion US dollars].

```{r}
gapminder %>%
  group_by(year, continent) %>%
  summarise(pop_in_bn = sum(pop) / 10^9) %>%
  ggplot(mapping = aes(x = year, y = pop_in_bn, fill = continent)) +
  geom_col()
    
    gapminder %>%
      group_by(year, continent) %>%
      summarise(gdp_in_tn = sum(pop * gdpPercap) / 10^12) %>%
      ggplot(mapping = aes(x = year, y = gdp_in_tn, fill = continent)) +
      geom_col()
```



m. Modify the figures from the previous question in such a way that they better emphasise the proportion of the total global population or total global GDP that different continents account for.
```{r}
  gapminder %>%
      group_by(year, continent) %>%
      summarise(pop_in_bn = sum(pop) / 10^9) %>%
      ggplot(mapping  = aes(x = year, y = pop_in_bn, fill = continent)) +
      geom_col(position = position_fill())
    
    gapminder %>%
      group_by(year, continent) %>%
      summarise(gdp_in_tn = sum(pop * gdpPercap) / 10^12) %>%
      ggplot(mapping = aes(x = year, y = gdp_in_tn, fill = continent)) +
      geom_col(position = position_fill()) 
    
```


n. Which position adjustment would you need to have the bars for each continent side-by-side rather than stacked on top of one another?

position_dodge

# Exercise 7.2

7.2 Exercise
According to the “Data visualisation with ggplot2” cheat sheet, both geom_col() and geom_bar() create bar plots.

a. Why does geom_col() require both the x and y aesthetic whereas geom_bar() only requires the x aesthetic?

geom_bar() only requires the x-aesthetic because it automatically counts the number of cars in each class and then plots the heights of the bars accordingly.

b. Recreate the following figure using geom_col() and geom_bar().

```{r}
# geom col 
mpg %>%
  count(class, name = "count") %>%
  ggplot()+
  geom_col(mapping = aes(x = class, y = count))
# you need to create a count for class before you can use it keep this in mind for next time 

# geom bar
mpg %>%
  ggplot() +
  geom_bar(mapping = aes(x = class))
# rather simple 
```

# Exercise 7.3

a.Recreate the following figure.

```{r}
mpg %>%
  mutate(year = as.factor(year)) %>% #turn them into factors and make your life easier
  ggplot(mapping = aes(x = class, y = cty)) +
  geom_boxplot(mapping = aes(colour = year))
  
```

b. Improve the following figure to deal with the overlap.
```{r}
ggplot(data = mpg) +
  geom_density(mapping = aes(x = cty, colour = class, fill = class), position = "stack",
               alpha = 0.1)
# better way is to increase the transparency
ggplot(data = mpg) +
      geom_density(
        mapping = aes(x = cty, colour = class, fill = class), 
        alpha = 0.1
      )
```

c. Read the documentation for geom_histogram() to learn how the number of bins/bin widths in a histogram can be controlled. Then explore different bin widths to improve data display in the following figure.

Bin width overrides bins and bin width is how wide you want the bins

```{r}
ggplot(data = mpg) +
  geom_histogram(mapping = aes(x = cty, fill = class))

# reading documentation frequency polygon has the same rules.
?geom_histogram

ggplot(data = mpg) +
  geom_freqpoly(mapping = aes(x = cty, colour = class), binwidth = 5)
```

# Exercise 7.4

a. Read again the documentation for geom_boxplot() to recall how the length of the whiskers in boxplots is determined (which in turn determines which points are treated as “outliers”).

```{r}
?geom_boxplot
```

Upper whisker goes from hinge to 1.5 of the interquartile range as a maximum, the lower whisker goes from hinge to the smaller value within 1.5 of the interquartile range of the hinge. Data beyond whiskers are considered outliers.

b. Recreate the following figure based on the filtered data for the year 2007 (colour and transparency level do not need to be reproduced exactly) which uses the rule that boxplots are automatically omitted for groups containing less than 20 observations.

```{r}
gapminder_2007 <- gapminder %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarise("gdpPercap"=sum(gdpPercap))

minimum <- 20

scatter <- gapminder_2007 %>%
  ggplot() +
  geom_point(mapping = aes(x = continent, y = gdpPercap), position = position_jitter(width = 0.1, height = 0)) +
   geom_boxplot(
             data          = . %>% filter(continent >= minimum),
             fill          = "#00B6EB",
             colour        = "#00B6EB",
             alpha         = 0.2,
             outlier.shape = NA,
             varwidth      = TRUE
           )

scatter


```
c. Modify the code from the previous question so that points designated as outliers or points in groups without sufficient numbers of observations are automatically labeled as shown below. *Note:* you can automatically extract the endpoints of the whiskers computed by `geom_boxplot()`. The easiest way to access these is from the data frame `ggplot_build(p)$data[[i]]` where `p` is the variable to which your figure has been assigned and assuming that the boxplots form the `i`th layer. Likewise, you can access the coordinates of the jittered points from the data frame `ggplot_build(p)$data[[j]]` assuming that the jittered points form the `j`th layer. Finally, the figure below also uses the package [ggrepel](https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html) which provides a modification of `geom_text()` called `geom_text_repel()` that deals better with overlapping labels.

```{r}
ggplot_build(p)$data[[1]] %>% # the jittered points form the first layer
      tibble() %>%
      select(c(x, y)) %>% 
      mutate(x = as.numeric(x)) ->
      points_info
    
    # Finds the continents for which we have drawn a boxplot.
    df %>%
      ungroup() %>%
      filter(continent_count >= n_min) %>%
      pull(continent) %>% # extracts only the column continent
      fct_drop() %>% # drops unused factor levels
      fct_unique() -> # returns only unique factor levels while maintaining the level ordering
      levels_with_boxplot
      
    ggplot_build(p)$data[[2]] %>% # the boxplots form the second layer
      tibble() %>%
      select(c(ymin, ymax, group)) %>%
      rename(continent = group) %>%
      mutate(continent = as.factor(continent)) %>%
      mutate(continent = factor(
                           continent,
                           labels = levels(levels_with_boxplot)
                         )
      ) ->
      boxplot_info
    
    df2 <- bind_cols(df, points_info)
    df2 <- left_join(df2, boxplot_info) %>%
      mutate(is_labeled = FALSE) %>%
      mutate(is_labeled = replace(is_labeled, y < ymin | y > ymax | continent_count < n_min, TRUE))
    
    library("ggrepel") 
    q <- df2 %>%
      filter(is_labeled == TRUE) %>%
      geom_text_repel(mapping = aes(x = x, y = y, label = country), size = 4)
    
    p + q
    ```
```{r}
geomco
```


