---
title: "practical_2"
output: html_notebook
---

# 2.0 Exercise

## Start a new R notebook named practical_02.Rmd and save it to the rmd folder of your RStudio project practicals. Make sure that all the R packages you need are loaded at the beginning. All the following exercises should be attempted within R code chunks of this notebook. (complete)

# 2.1 Exercise (complete)
## Create a variable that is of type:

a. double, 

```{r}
a <- 5.5

typeof(a)
```

b. integer,

```{r}
b <- 65L

typeof(b)
```

c. character,

```{r}
c <-  "hello world"

typeof(c)
```

d. logical.

```{r}
d <- TRUE

typeof(d)
```

e. Set

```{r}
x <- TRUE
y <- "TRUE"
a <- 1
b <- 1L
c <- "1"
```

Without executing the code, state which of these operations are allowed. For those that are, state what the result is and its data type. (complete )

```{r}
ax <- a - x # 0 (dbl)
xb <- x + b # 2 (int)
cx <- c * x # will cause error
xy <- c(x, y) # character type it will coerce to character type ("TRUE","TRUE")
ab <- c(a, b) # (1,1) will coerce to double
abc <- c(a, b, c) # ("1","1","1") will coerce to character


```

# 2.3 Exercise

Read about the function paste() in the documentation and use it to write:

`"pi is approximately equal to 3.14159265358979"`

to the console but without typing in any numbers.



```{r}
?paste

paste("pi is approximately equal to", pi)

```

# 2.3 Exercise

a. Write code that assigns some real number to a variable x and then prints a message stating whether x is divisible by 2 or not. Hint: You may find the operator %% useful which gives the remainder term. That is, x %% 2 returns 0 whenever x is divisible by 2. (complete)


```{r}
# Using ifelse function

x <- 29

divisible_by_2 <- ifelse(
  test = x %% 2 == 0,
  yes = "divisible by 2",
  no = "not divisible by 2"
)

print(divisible_by_2)


```

b. Modify the code from the previous part to state whether x is divisible by 2, by 3 or by neither. Hint: x %% 3 returns 0 whenever x is divisible by 3. (complete)

```{r}
if (x %% 3 == 0 & x %% 2 == 0) {
  print("divisible by both")
} else if (x %% 3 == 0) {
  print("divisible by 3")
} else if (x %% 2 == 0) {
  print("divisible by 2")
} else {
  print("neither")
}
```

2.4 Exercise

a. Write a function normalise which takes a numeric vector x as an argument, subtracts the mean and then divides by the standard deviation of its elements, returning the result (hint: you may find the functions mean() and sd() helpful). (complete)

```{r}
normalise <- function(x) {
  average_by_standard_dev <-  mean(x) / sd(x)
  return(average_by_standard_dev)
}

x <-  c(1, 2, 3, 4, 5, 6, 9, 7)

normalise(x)
# check
mean(x) / sd(x)

```

b. Test your function from (a) on the vector x <- c(0, 2, -5, 1, 2) (you can verify the output of your code by y <- normalise(x) and then verifying that mean(y) is 0 and sd(y) is 1). (complete)

```{r}
x <- c(0, 2, -5, 1, 2)

normalise(x)
# check

mean(x) / sd(x)
```


c. What happens if you test your function from (a) on x <- c(0, NA, -5, 1, 2) and why? Modify your function so that it can be used to only normalise the non-missing values in a vector if explicitly requested by the user (i.e. the default behaviour of normalise(x) should remain unchanged). (incomplete)

```{r}
x <- c(0, NA, -5, 1, 2)

normalise(x)

# You get NA because of the NA
# modifying function
normalise <- function(x, na.rm = FALSE) {
  average_by_standard_dev <- (mean(x, na.rm = TRUE) / sd(x, na.rm = TRUE))
  return(average_by_standard_dev)
}

normalise(x, na.rm = TRUE)


# go back to this one completed it but how i did that confuses me

```

# 2.5 Exercise

Create a vector filled with 5, 4, 3, 2, 1 (in this order):

a. manually using the c() function,(complete)

```{r}
c(5, 3, 3, 2, 1)
```

b. using the : operator, (complete)

```{r}
c(5:1)
```


c. using seq() with a specified step size, (complete)

```{r}
seq(from = 5, to = 1, by = -1)
```

d. using seq() with a specified vector length. (complete)

```{r}
seq(from = 5, to = 1, length.out = 5)
```

# 2.6 Exercise

a. Recreate the vector c(1:3, 1:3) using rep() with times argument,(complete)

```{r}
one_two_three <- c(1, 2, 3)
rep(one_two_three, times = 2)
```


Recreate the vector c(1, 1, 2, 2, 3, 3) using rep() with each argument. (complete)

```{r}
rep(one_two_three, each = 2)
```

# 2.7 Exercise

R provides a built-in constant vector letters which is as follows.

```{r}
letters
```

a. Construct a numeric vector x such that typing letters[x] into the console yields the given output. For each line of output, type in no more than 2 numbers to define x. (complete)

`[1] "a"`
`[1] "c" "d" "e"`
`[1] "e" "d" "c"`
`[1] "b" "d" "f" "h" "j" "l" "n" "p" "r" "t" "v" "x" "z"`
`[1] "z" "w" "t" "q" "n" "k" "h" "e" "b"`

```{r}
# a
x <- 1
letters[x]
# c d e
x <- c(3:5)
letters[x]
# e d c
x <- c(5:3)
letters[x]
# b d f h j l n p r t v x z
x <- seq(from = 2, to = length(letters), by = 2)
letters[x]
# z w t q n k h e b
x <- seq(from = length(letters), to = 2, by = -3)
letters[x]

```

b.Without running the code, predict the output each of the following eight chunks of code. (complete)

```{r}
x <- letters
x[length(x)] <- "a"
x # last letter will be replaced by "a"

x <- letters
x[1:length(x)] <- "bla"
x # you will get bla 26 times

x <- letters
x[1:length(x)] <- 0
x # you will get 0 26 timws 

x <- letters
x <- "a"
x # x will just be a, a will be stored as x instead 

x <- letters
y <- 1:length(letters)
x[y > 10] <- "bla" 
x # any letter past j will be replaced by "bla"

c("b", "c") %in% letters # Will return TRUE if letters appear in vector 

which(letters == "c") # will tell index position of c

letters[27] # should be NA or error.
```


c. Without running the code, predict the output each of the following three lines of code.

```{r}
intersect(letters[2], letters) # b

setdiff(letters, letters[3:length(letters)]) # a b

letters[which.min(c(2, 1, 2))] # do not understand the third one.
```

# 2.8 Exercise

Write a function count_na() that takes a vector as an argument and then returns the number of missing values in the vector.

a. using a for-loop,

```{r}
x <- c(2, NA, 4, NA, 9, 10, 13, NA)

count_na <- function(x) {
  n_na <- 0
  for (na in x) {
    if (is.na(na)) {
      n_na <-  n_na + 1
    }
  }
  return(n_na)
}

count_na(x)

# come back to this later... you suck at for loops
```

b. without using any loop.
```{r}
count_na <- function(x) {
  n_na <- sum(is.na(x))
  return(n_na)
}

x <- c(1, NA, NA, 7, 4, 8)
count_na(x)
```

# 2.9

In the lecture notes, we have learned that we should use the operators && and || for single values and & and | for vectors. What does the following expression do and why?


```{r}
c(TRUE, TRUE) && c(TRUE, FALSE)
```

It is a logical operator which checks if both conditions are true, this will return True, however with the vectorised operator it will return false. 
The "double" logical binary operators **&&** and **||** only consider the **first** **element** of the vectors on each side, i.e. this expression evaluates.

# 2.10 Exercise

Go through all the code written for the previous answers and make sure that it follows the style guidelines from Chapter 2.

I will use lintr.

