---
title: "My Practicals"
subtitle: "Practical_1"
author: "Sam KD"

output: 
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r echo=FALSE, results='hide',message=FALSE}
library("tidyverse")
```

```{r setup,results='hold'}

```

# Exercises

## a. Create a new RStudio project called practicals. (completed)

## b. Create subfolders data, rmd, and figures in your project folder.(completed)

## c. Create a new R notebook and save it to the subfolder rmd. Change the title, include a subtitle and your name as the author. Make sure that all the R packages you need are loaded at the beginning of your notebook. In your notebook, include:

1. numbered sections, subsections and sub subsections,(complete)
2. a table of contents,(complete)
3. a list whose first level is numbered but whose second level is not,
bold and italic text, (complete)
verbatim code, (complete)
text with superscripts, (complete)
a reproduction of the following text (including mathematical notation and including the link) from the “General linear model” section of the Wikipedia article on regression analysis. Hint: if you don’t know the LaTeX/mathjax syntax for all the symbols shown here, you can simply copy-and-paste the equations from the original URL. (copy from stats) (complete)

$$
yi=\beta_1x_i+\beta_2 x_i2 + ... + \beta_p xi_p +\epsilon i,
$$
# 
1. list 1

+ list 2:
**bold**
*italic*

### 

`verbatim code`

### 

text^sub^

# Exercise

## Include versions of the following R code chunk in your notebook created in Exercise 1:

```{r}
x <- 42
x
log(-1)
```

Then modify the chunk options so that the knitted document,

1. executes the code and shows code, output and warnings below the full code chunk, (complete)
```{r}
x <- 42
x
log(-1)
```
2. executes the code and shows code, output and warnings but doesn’t split up the code chunk so that code and output are each displayed in a single box. (complete)
```{r results='hold'}
x <- 42
x
log(-1)
```

3. executes the code and shows code and output below the full code chunk but no warnings,(complete)
```{r warning=FALSE}
x <- 42
x
log(-1)
```
4. executes the code and shows code but no output and no warnings, (complete)
```{r warning=FALSE,results='hide'}
x <- 42
x
log(-1)
```
5. executes the code but shows no code, no output and no warnings, (complete)
```{r echo=FALSE,warning=FALSE,results='hide'}
x <- 42
x
log(-1)
```
6. does not execute the code but shows the code,
```{r eval=FALSE}
x <- 42
x
log(-1)
```

7. neither shows nor executes the code. (complete)
```{r eval=FALSE,echo=FALSE}
x <- 42
x
log(-1)
```

# exercise

##

a. Include a hidden (but still executed) code chunk at the start of your notebook which loads the tidyverse and here packages. (complete)

b. Add a setup chunk at the start of your notebook which globally sets results = "hold" but which should not be displayed in the rendered document and whose output should not be visible therein. (complete)

c. Download the image cat.jpg and save it to the figures subfolder of your project. Include the image in your notebook using the absolute file path. Similarly, download the data set mcdonalds_menu.csv and save it to the data folder of your project. Include a code chunk in your notebook which reads the data set using an absolute file path, i.e. `read_csv(<PATH>)` (complete)

[cat](../figures/cat.jpg)

```{r}
read.csv("~/Documents/loughborough_work/introduction_to_data_science/intro_to_r_project/practicals/my_practicals/data/mcdonalds_menu.csv")
```

ii. Where you need to replace `<PATH>` by the relevant absolute path to the data set. Then knit the document. Make sure that the image and the data set appears in the rendered document before moving to the next question. (does not work for me lol)

Move the project folder (with all its contents) to another location on your computer and knit it again. What happens and why? Find a way to fix the problem.

It will not knit to fix you can set the new place as the working directory 

# Exercise

1. Add the a chunk with the following R code chunk to your notebook. You don’t need to understand the details of what rnorm()does. All that matters is that it generates a collection of (pseudo-)random numbers1 and mean(x) then computes the average of those numbers.

```{r}
x <- rnorm(n = 10)
y <- print(round(mean(x), digits = 2))
```

a. Execute the code chunk. What is the output? Execute the chunk again. Explain why the output has changed.(complete)

It creates random numbers so it will always change

b. Add the sentence “The mean of x <RESULT>.” below your code chunk where “<RESULT>” should automatically take the same value as the one printed to the console even if you re-knit the document or execute the chunk again. Remember that <RESULT> is meant to b placeholder for the result (i.e. you are not meant to type in any < or > symbol). (completed)

The mean of x is `r y `




