---
title: "practical_4"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
```

# 4.1 Exercise (complete)

Consider the starwars data set from the dplyr package.
```{r}
starwars
```

a. Use the function glimpse() to get an overview of which variables this data set contains and what data types the variables have.

```{r}
glimpse(starwars)
```


b. Find as many different ways as you can to output Yoda’s height (you may use that Yoda entry is the 19th row in starwars). The output should always be an atomic vector (i.e. not a list or a data frame). Achieve this:

```{r}
# matrix indexing
starwars[[19,2]]
#integer indexing
starwars[[2]][19]
# by name indexing
starwars[["height"]][19]
# by $ operator
starwars$height[19]
# by pull and filter 
starwars %>%
  filter(name == "Yoda") %>%
  pull(height)
```

i . by interpreting the data frame as a matrix,(complete)

```{r}
# matrix indexing
starwars[[19,2]]
```

ii. by interpreting the data frame as a list,(complete)

```{r}
#integer indexing
starwars[[2]][19]
# by name indexing
starwars[["height"]][19]
# by $ operator
starwars$height[19]
```

iii. by using dplyr verbs. (complete)

```{r}
# by pull and filter 
starwars %>%
  filter(name == "Yoda") %>%
  pull(height)
```


# 4.2 Exercise (needed solutions, complete though)

Consider again the starwars data set. Write R code that achieves the following tasks. For each task, the code should use a single pipe, (multiple %>% operators are allowed) starting with the original starwars data set, i.e. the beginning should always read starwars %>% (except in (j) where you might need to define some other variables first).

a. Order the cases (rows) set alphabetically according to the character name.

```{r}
starwars %>%
  arrange(name)
```

b. Order the cases first by species and by character name (alphabetically) within each species.

```{r}
starwars %>%
  arrange(species,name) 
```

c. Count the number of characters whose homeworld is Tatooine or not recorded in the data set. The code should produce a single number (not a data frame!).

```{r}
starwars %>%
  filter(homeworld == 'tattoine' | is.na(homeworld)) %>%
  count() %>%
  pull()
  
```

d. Create a table which only includes the variable name and variables whose names include the string “color” without using the full column names or column positions.

```{r}
starwars %>%
  select(name, contains("color"))
```

e. Create a table which gives the average height [in cm] of female characters for each species. Don’t forget to ensure that all variables have appropriate names.

```{r}
starwars %>%
  filter(sex == "female") %>%
  group_by(species) %>%
  summarise("female avg height (cm)" = mean(height,na.rm = TRUE))
```

f. Create a table which counts the proportions of different eye colours within each species. Don’t forget to ensure that all variables have appropriate names.

```{r}

starwars %>%
  group_by(species, eye_color) %>%
  count() %>%
  group_by(species) %>%
  mutate(proportion = n/sum(n)) %>%
  select(!n)

```


g. Display the range (difference between maximum and minimum) of the height of the characters within each species.

```{r}
starwars %>% 
      group_by(species) %>%
      summarise(min_height = min(height, na.rm = FALSE), max_height = max(height, na.rm = FALSE)) %>%
      mutate(range_height = max_height - min_height,) %>%
      select(-c(min_height, max_height))
```

h. Add a variable rel_height to the data frame which is such that rel_height represents the height of a character relative to

i. the mean of the heights of all characters in the data set,

```{r}
   starwars %>%
      select(name, height, species) %>%
      mutate(rel_height = height / mean(height, na.rm = TRUE))
```

the mean of heights of all characters of the same species in the data set.

```{r}
starwars %>%
      select(name, mass, species) %>%
      group_by(species) %>%
      mutate(rel_height = mass / mean(mass, na.rm = TRUE))
```

i. Without typing in any column names,

i. create a new data frame which only contains the columns which are type character or type list.

```{r}
starwars %>%
  select(where(is.character), where(is.list))
```

ii. change all the columns which are character (except the column name) to factor.

```{r}
starwars %>%
  mutate(across(where(is.character) & !name, as.factor))
```

j. Create a data frame which gives the number of characters whose height is “very short” (height≤100), “short” (100<height≤170), “average” (170<height≤200) or “tall” (height>200),


i. using the case_when() function,

```{r}
 breaks <- c(0, 100, 170, 200, Inf)
    labels <- c("very short", "short", "average", "tall")
    
    # solution using case_when()
    starwars %>% 
      mutate(height_category = case_when(
          is.na(height) ~ NA_character_, # each data type has its own version of NA and here we need to use the right one otherwise we'll get an error
          height <= breaks[2] ~ labels[1],
          height <= breaks[3] ~ labels[2],
          height <= breaks[4] ~ labels[3],
          TRUE ~ labels[4]
        )
      ) %>%
      group_by(height_category) %>%
      summarise(no_of_characters = n())
```


ii. using the base R function cut() (which you may need to look up).

```{r}
starwars %>% 
      group_by(height_group = cut(height, breaks = breaks, labels = labels)) %>%
      summarise(no_of_characters = n()) 
```

# 4.3 Exercise 
 
Download the files cabi_stations.csv as well as cabi_rides_201901.csv and cabi_rides_201902.csv. These files contain collected by Capital Bikeshare (CaBi) – a publicly-owned bicycle sharing system that serves areas in and around Washington DC in the United States. Specifically, the file cabi_stations.csv contains information about the pick-up and return stations1; the files cabi_rides_201901.csv and cabi_rides_201902.csv contain information about rides undertaken by CaBi users in January and February 20192.

Include the following code chunk in your notebook. We will learn more about the function read_csv() in Chapter 6. For the moment, all you need to know is that it reads the content of (CSV) data file into a variable in R (ignore the output under Column specification ... shown in the console when executing these lines).

```{r}
stations <- read_csv(here("data", "cabi_stations.csv"))
trips_01 <- read_csv(here("data", "cabi_trips_201901.csv"))
trips_02 <- read_csv(here("data", "cabi_trips_201902.csv"))
```

a. Use the function glimpse() to get an overview of which variables these data sets contain and what data types the variables have.

```{r}
glimpse(stations)

glimpse(trips_01)
glimpse(trips_02)
```

b. combine trips_01 and trips_02 into a single tibble called trips.

```{r}
trips <- trips_01 %>% bind_rows(trips_02)
```

c. Add columns start_longitude, start_latitude, end_longitude and end_latitude station to trips. These should contain information about coordinates of the start and end of each ride. No other columns should be added.

```{r}
trips
stations

  trips <- trips %>% 
      left_join(stations, by = c("start_station" = "address")) %>%
      rename(start_longitude = longitude, start_latitude = latitude) %>%
      left_join(stations, by = c("end_station" = "address")) %>%
      rename(end_longitude = longitude, end_latitude = latitude) %>%
      select(colnames(trips) | contains("longitude") | contains("latitude"))
```




