---
title: "practical_3"
output: html_notebook
---

# 3.1 Exercise (complete)

+ Create a (4,2)-matrix A whose first row is filled with 1s, whose second row is filled with 2s and whose third row is filled with 3s and so on:

a. by using the function matrix(),

```{r}
matrix(1:4, nrow =4, ncol = 2)
```

b. by using the function matrix(..., byrow = TRUE) (Hint: remember the function rep() from Practical 2),

```{r}
# not sure about this one chief look up documentation and solutions 
?matrix
matrix(rep(1:4, each = 2), ncol=2, byrow = TRUE)
```

c. by binding together two vectors by column,

```{r}
a <- c(1:4) 
b <- c(1:4)

cbind(a,b)
```


d. by binding together four vectors by row.

```{r}
a <- c(1,1)
b <- c(2,2)
c <- c(3,3)
d <- c(4,4)

rbind(a,b,c,d)

```

Define the following function and numeric vectors.


# 3.2 Exercise (complete)

```{r}
multiply <- function(x, y) {
  return(x * y)
}

x <- 1:3
y <- 3:1
```

a. Store multiply, x and y in an unnamed named list called l.

```{r}
l <- list(multiply, x, y )

```

b. Assign names “fun”, “x” and “y” to the three list elements.

```{r}
names(l) <- c("fun", "x","y" )

l
```

c. Create the list l again but this time assign the names already during list creation.

```{r}
l <- list("fun"=multiply,"x"=x, "y"= y)

l
```

d. Apply the function multiply to the vectors x and y using only the list l.

```{r}
l$fun(l$x, l$y)
# or 
l[[1]](l[[2]],l[[3]])
# or from solutions
l[["fun"]](l[["x"]], l[["y"]])
```

e. A list can even be included as an element in another list. Set:

```{r}
w <- list("l" = l)
```

i. Without executing any command, state how many elements are in w. Verify your answer using length().

There is only 1 element in w, the as it contains the list only

```{r}
length(w)
```

ii. Write at least six different expressions which all access the numeric vector y that is stored within w.


```{r}
w
# by name
w$l$y
# by name character
w[["l"]][["y"]]
# by integer indexing
w[[1]][[3]]
# logical indexing?
w[[TRUE]][[FALSE, FALSE, TRUE]]
# doesn't work
# indexing and name
w[[1]]$y
w$l[[3]]
# character name and name
w[["l"]]$y

```



