---
title: "practical_5"
output: html_notebook
---

```{r}
library(lubridate)
library(tidyverse)
```

# 5.1 Exercise (completed, needed solutions so you are weak on factors)

Include the following code chunk in your notebook and execute it.

```{r}
films <- c(
  "A New Hope", "The Empire Strikes Back", "Return of the Jedi",  
  "The Phantom Menace", "Attack of the Clones", "Revenge of the Sith", 
  "The Force Awakens", "The Last Jedi", "The Rise of Skywalker"
)
```

Assume that we have asked five classmates to pick their favourite Star Wars movie (only those listed in films should have been an option). We have recorded their answers in the vector survey_results_2 as follows.

```{r}
survey_results_2 <- 
  c("Rogue One", "A New Hope", "The Force Awakens", "A New Hope", "The Empire Strikes Back")
```

a. Create a factor with these modified survey results and levels = films. What happens to the first data point in this case?

```{r}

survey_results_2 %>%
  factor(levels = films)
```

The first data point appears as an NA because it is not a level on there.


b. Let us assume that “Rogue One” was indeed an option in the survey. Create a factor z which takes this into account

```{r}
# much quicker than manually adding a factor using c() then using factor again
z <- factor(survey_results_2, levels = c(films, "Rogue One"))
z
```

c. Reorder the levels of z so that the order reflects the Star Wars chronological order (i.e. Episode I, Episode II, Episode III, Rogue One, Episode IV, …, Episode IX).

```{r}
z <- factor(survey_results_2, levels = c(films[4:6], "Rogue One", films[c(1:3, 7:9)]))
z
```

d. Ensure that we can easily perform comparisons between the elements of z.

```{r}
ordered_z <- ordered(z, levels = levels(z))

# to check
ordered_z[1] < ordered_z[2] 

```


# 5.2 Exercise
Consider the following data set of (abbreviated) month names.

```{r}
months <- c("Feb", "March", "DEC", "Nov", "Mar", "Jul", "Jan", "Jul", "jul")
```


a. Convert months into a factor z whose levels are the names of all twelve months plus any other variation/misspelling of a month name found in the vector months. (Tip: The constants month.abb and month.full available in R contain the (abbreviated) names of the twelve months of the year).

```{r}
# use unique to remove duplicate levels, because duplicate levels cause an error
z <- months%>%
  factor(levels = unique(c(month.abb, months)))

```


b. Clean up the factor z to correct for the variations/misspellings of some of the months,

i. without using forcats functions,

```{r}
 levels(z) <- c(month.abb, "Mar", "Dec", "Jul")
```


ii. using the package forcats.

```{r}
z %>% fct_collapse(
  "Mar" = "March",
  "Jul" = "jul",
  "Dec" = "DEC"
 )
```


c. Consider the following character vector.

```{r}
months <- c("January", "December", "March", "March", "December")
```

Use the fact that a factors stores its values internally as integers to convert months into the numeric representation c(1, 12, 3, 3, 12).

```{r}
unclass(factor(months, levels = month.name))
```

# 5.3 Exercise
Without manually calculating, find out how many seconds have passed between now and the time and day you were born (if you don’t know the time-of-day, pick 00:00:00 but make sure to use the right time zone).

```{r}
bday <- ymd_hms("1999-09-28 04:00:56") - now()

bday_seconds <- as.duration(bday)

paste(bday_seconds, "have passed since my birth")
```

# 5.4 Exercise (completed)

You are living in the UK and you have arranged to talk to your parents in Nairobi, Kenya, via Skype/Zoom on 2021-10-10 at 3 pm. But you forgot whether the clock time was meant to be in your time zone or in their time zone. Calculate

a. the clock time in your time zone if the arranged time referred to the clock time in your parents’ time zone,

```{r}
talk_with_parents <- ymd_h("2021-10-10 15", tz = "Africa/Nairobi")

talk_with_parents

talk_with_parents_utc <- talk_with_parents %>%
  with_tz(tzone =) # defaults to UTC

talk_with_parents_utc
```


the clock time in your parents’ time zone if the arranged time referred to the clock time in your time zone.


```{r}
talk_with_parents_2 <- ymd_h("2021-10-10 15")

talk_with_parents_kenya <- talk_with_parents_2 %>%
  with_tz(tzone = "Africa/Nairobi")

talk_with_parents_kenya
```


# 5.5 Exercise

Consider again the starwars data set.

```{r}
starwars
```

a. Return a table which contains only the data for those characters whose name includes a “-”. 

```{r}
starwars %>%
  filter(str_detect(name, "-")) %>%
  select(name)
```


b. Return a table in which the names of all characters are abbreviated to the first three letters/numbers.

```{r}
# my method lol
str_sub(string = starwars$name, start = 1L, end = 3L ) %>%
  as.tibble() %>%
  rename("name" = value)

# proper way

starwars %>%
  mutate(name = str_sub(name, start = 1L, end = 3L))
```
c. Return a table in which each “2” in the name of a character is replaced by “Two”. (complete,needed solutions)

```{r}
starwars %>% 
  mutate(name = str_replace_all(name, pattern = "2", replacement = "Two"))
```


# 5.6 Exercise

a. Create a tibble with twelve columns, each filled in such a way that the tibble ends up with all twelve different data types mentioned in the table in Section 1.3.2 of Chapter 4. (Only got to 10 without solutions)

```{r}
# you can use previous columns to create new types

tbl <- tibble(
  a <- c(TRUE,FALSE,TRUE),
  b <- c(1L, 2L, 3L),
  c <- c(1, 2, 3),
  d <- c("one", "two", "three"),
  e <- list(TRUE,1,"one"),
  f <- list("x" = TRUE, "y" = 3, "z" = "why"),
  g <- factor(c("a","b","c"),levels = "a","b","c", ordered = FALSE),
  h <- factor(c("a","b","c"),levels = "a","b","c", ordered = TRUE),
  i <- as_date(c("2021-01-01","2020-01-01","2019-01-01")),
  j <- c(ymd_h("2021-01-02 04"), ymd_hm("2020-07-17 15:27"), ymd_hms("2019-09-28 18:30:30", tz = "Africa/Accra")),
  k = c(today() - (i - 1)),
  l = as.duration(k)
)

```

```{r}
glimpse(tbl)
```

