---
title: "practical_6"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
```

# 6.1 Exercise
Parse the following strings using a suitable parse_*() function. (slightly more difficult than expected)

```{r}
"1,5" # to be parsed as 1.5

parse_double("1.5",locale = locale(decimal_mark = ","))

"September 21, 2020 16:25:59" 

parse_datetime("September 21, 2020 16:25:59 UTC", format = "%B %d, %Y %T %Z")

"€1.999,99" # the symbol makes it parse number instead

parse_number("€1.999,99",locale = locale(grouping_mark = ".", decimal_mark = ","))
```

# 6.2 Exercise
Download the file ramen_ratings.csv. This data set contains contains ratings of different types of ramen and is a modified version of a data set on kaggle (but use the file from Learn and not the one from kaggle)

a. Inspect the structure of the CSV file by printing the first four lines to the console.

```{r}
read_lines(here("data","ramen_ratings.csv"), n_max = 4)
```
csv 2 delim deffo

b. Read the data into R.

```{r}
ramen <- read_csv2(here("data","ramen_ratings.csv"), skip = 1)
```


c. Inspect the tibble created in Question 1. What are the data types of the columns? Why are the names of the columns Review # and Top Ten encapsulated in back ticks? Reformat the column names to snake_case.

The data types in of the columns are all character except review which is a double. Review # and Top Ten have back ticks because they are columns with spaces in them.

```{r}
# inspecting tibble
glimpse(ramen)
colnames(ramen)

# reformatting variables.
ramen <- clean_names(ramen)
```


c. Above, we learned that R uses heuristics to guess a sensible data type for each column. Why is the column stars of type character and not double?

R is confused because of the "unrated" values so it made everything a character.

d.Convert Stars to a double vector using a suitable parser. (completed but forgot mutate function, don't forget next time.)

```{r}
# when making conversions do not forget mutate
ramen <- ramen %>%
  mutate(stars = parse_double(stars, locale = locale(decimal_mark = ",")))

ramen <- ramen %>%
  mutate(review_number = as.integer(review_number))
```


e. Change the data type of stars. What do the parsing failures mean for your analysis.

parsing failures shows that there are some observations that are incomplete and we may have to remove them depending on how important the rest of the information is.


# 6.3 Exercise
Download the file market_cap.csv. This data set contains time-series data on the market capitalisation of ten US companies. Load the data into R and shape it into a tidy format (don’t forget to ensure that all columns have sensible data types).

```{r}
# always read lines

read_lines(here("data","market_cap.csv"), n_max = 4)

#clearly read csv

```

```{r}
market_cap <- read.csv(here("data","market_cap.csv"))

market_cap
```
not difficult make the first set of columns years and the values market cap (pivot longer) (complete)

```{r}
market_cap <- market_cap %>%
  pivot_longer(
    cols      = -company, # use all columns except "company"
    names_to  = c("year") 
  ) 

market_cap <- market_cap %>%
  mutate(year = str_remove(year,"X"))

market_cap %>% 
  mutate(year = as_date(year))
```

# 6.4 Exercise
The following code recreates the non-tidy data frame for the cats-and-dogs example from Section 2.1 in Chapter 6.

```{r}
df_nontidy <- tibble(
  ` ` = c("selfless", "selfish"), 
  cats = c(4L, 19L), 
  dogs = c(25L, 3L)
)
df_nontidy
```

a. Reshape the data into the same tidy format shown in Section 2.1 in Chapter 6. Don’t forget to ensure that all columns have sensible data types.

```{r}
    df_tidy <- 
      df_nontidy %>%
      pivot_longer(
        cols = c(cats, dogs), 
        names_to = "animal", 
        values_to = "count"
      ) %>%
      rename(character = ` `)
    
    df_tidy
```


b. Reverse the operation so that you obtain the original non-tidy data frame df. 
```{r}
df_nontidy <-
      df_tidy %>%
      pivot_wider(names_from = animal, values_from = count) %>%
      rename(` ` = character, dogs = dogs, cats = cats)

df_nontidy
    
```

6.5 Exercise
Perform the reverse of the operations shown in Section 2.2 of Chapter 6, i.e. recreate the non-tidy tables df_nontidy_1 and df_nontidy_2 shown in Section 2.2. by lengthening/widening the tidy table df_tidy.


```{r}
df_tidy <- table1 %>%
  mutate(
    country = str_replace_all(country, c("Afghanistan" = "A", 
                                         "Brazil" = "B", 
                                         "China" = "C"))
  ) %>%
  mutate(across(c(cases, population), round, -3)) %>%
  mutate(across(c(cases, population), as.integer))

df_tidy
```
```{r}
df_nontidy_1 <- df_tidy %>%
  pivot_longer(
    cols     = c(cases, population), 
    names_to = "type", 
    values_to = "count"
  )

df_nontidy_2 <- df_tidy %>%
  pivot_wider(
    names_from  = year, 
    values_from = c(cases, population)
  )
```

# 6.7 Exercise

Use drop_na() to perform the reverse of the operation from Section 2.4 of Chapter 6, i.e. to convert from explicit missingness (df_missing_1) to implicit missigness (df_missing_2), where df_missing_1 is obtained as follows.


```{r}
df_missing_1 <- df_tidy
df_missing_1[2, ] <- NA
```

```{r}
df_missing_1[2, ] <- NA
```

# for 6.8 grab from solutions 


